---
title: "Pareto"
author: "Ulrich Riegel"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The (European) Pareto distribution is probably the most popular distribution for modelling large losses in reinsurance pricing. There are good reasons for this popularity, which are discussed in detail in Fackler (2013). We recommend  Philbrick (1985)  and Schmutz *et.al.* (1998) for an impression of how the (European) Pareto distribution is applied in practice. 

In cases where the Pareto distribuion is not flexible enough, pricing actuaries sometimes use picewise Pareto distribuions. For instance, a Pareto alpha of 1.5 is used to model claim sizes between USD 1M and USD 5M and an alpha of 2.5 is used above USD 2.5M. Riegel (2018) provides an algorithm for matching a tower of layer losses with a collective risk model using a Piecewise Pareto distribution.

The main purpose of the Pareto package is to provide this matching algorithm. In addition the package provides various tools for the Pareto and the Piecewise Pareto distribution, which are useful for pricing of reinsurance treaties.

## Pareto distribution

**Definition:**
Let $t>0$ and $\alpha>0$. The *Pareto distribution* $\text{Pareto}(t,\alpha)$ is defined by the distribution function $$
F_{t,\alpha}(x):=\begin{cases}
0 & \text{ for $x\le t$} \\
\displaystyle 1-\left(\frac{t}{x}\right)^{\alpha} & \text{ for $x>t$.}
\end{cases}
$$
This version of the Pareto distribution is also known as *Pareto type I*, *European Pareto* or *single-parameter Pareto*.

Why is the Pareto distribution so popular?

**Lemma:**

* Let $X \sim \text{Pareto}(t,\alpha )$. Then $cX \sim \text{Pareto}(ct,\alpha )$ for all $c>0$.
* Let $X \sim \text{Pareto}(t_{1} ,\alpha )$. For $t_2 > t_1$ we then have $X|(X>t_2 ) \sim \text{Pareto}(t_2 ,\alpha )$

**Consequences:**

* The *Pareto alpha* is invariant wrt scaling (which implies that $\alpha$ does not depend on currencies and inflation) \pause
* For Pareto distributed data the Pareto alpha does not depend on the reporting threshold \pause
* For layers and thresholds above $t$ the ratio between expected layer losses and/or excess frequencies  depends only on $\alpha$ (and not on   $t$)

### Layer mean:

Let $X\sim \text{Pareto}(t,\alpha)$ and $a, c\ge 0$. Then
$$
E(\min[c,\max(X-a,0)]) = \int_a^{c+a}(1-F_{t,\alpha}(x))\, dx. =: I_{t,\alpha}^{\text{$c$ xs $a$}}
$$
is the layer mean of $c$ xs $a$, i.e. the expected loss to the layer given a single loss  $X$.

*Example:* $t=500$, $\alpha = 2$, Layer 4000 xs 1000
```{r}
library(Pareto)
Pareto_Layer_Mean(4000, 1000, 2, t = 500)
```

### Pareto extrapolation

Consider two layers $c_i$ xs $a_i$ and a $\text{Pareto}(t,\alpha)$ distributed severity with sufficiently small $t$. What is the expected loss of $c_2$ xs $a_2$ given the expected loss of $c_1$ xs $a_1$? 

*Example:* Assume $\alpha = 2$ and the expected loss of 4000 xs 1000 is 500. Calculate the expected loss of the layer 5000 xs 5000.

```{r}
Pareto_Extrapolation(4000, 1000, 5000, 5000, 2) * 500
Pareto_Extrapolation(4000, 1000, 5000, 5000, 2, ExpLoss_1 = 500)
```


### Pareto alpha between two layers:

Given the expected losses of two layers, there is typically a unique Pareto alpha $\alpha$ which is consistent with the ratio of the expected layer losses. 

*Example:* Expected loss of 4000 xs 1000 is 500. Expected loss of 5000 xs 5000 is 62.5. Alpha between the two layers:

```{r}
Pareto_Find_Alpha_btw_Layers(4000, 1000, 500, 5000, 5000, 62.5)
```
Check: see previous example


### Pareto alpha between a frequency and layer:

Given the expected excess frequency at a threshold and the expected loss of a layer, then there is typically a unique Pareto alpha $\alpha$ which is consistent with this data.

*Example:* Expected frequency in excess of 500 is 2.5. Expected loss of 4000 xs 1000 is 500.  \pause Alpha between the frequency and the layer:
```{r}
Pareto_Find_Alpha_btw_FQ_Layer(500, 2.5, 4000, 1000, 500)
```
Check:
```{r}
Pareto_Layer_Mean(4000, 1000, 2, t = 500) * 2.5
```

### Matching the expected losses of two layers:

Given the expected losses of two layers, we can use these techniques to obtain a Poisson-Pareto model which matches the expected loss of both layers.

*Example:* Expected loss of 30 xs 10 is 26.66 (Burning Cost).  Expected loss of 60 xs 40 is 15.95 (Exposure model).
```{r}
Pareto_Find_Alpha_btw_Layers(30, 10, 26.66, 60, 40, 15.95)
```
Frequency @ 10:
```{r}
26.66 / Pareto_Layer_Mean(30, 10, 1.086263)
```
A collective model $\sum_{n=1}^NX_n$ with $X_N\sim \text{Pareto}(10, 1.09)$ and $N\sim \text{Poisson}(2.04)$ matches both expected layer losses.


### Frequency extrapolation and alpha between frequencies:}

Given the frequency $f_1$ in excess of $t_1$ then the frequency $f_2$ in excess of $t_2$ can directly be calculated as follows:
$$
f_2 = f_1 \cdot \left(\frac{t_1}{t_2}\right)^\alpha
$$
Vice versa, we can calculate the Pareto alpha, if the two excess frequencies $f_1$ and $f_2$ are given:
$$
\alpha = \frac{\log(f_2/f_1)}{\log(t_1/t_2)}.
$$

*Example:*

Expected frequency excess 1000 is 2. What is the expected frequency excess 4000 if we have a Pareto alpha of 2.5?
```{r}
t_1 <- 1000
f_1 <- 2
t_2 <- 4000
(f_2 <- f_1 * (t_1 / t_2)^2.5)
```
Vice versa:
```{r}
Pareto_Find_Alpha_btw_FQs(t_1, f_1, t_2, f_2)
```


add LayerVariance, Simulation etc


## Piecewise Pareto distribution

**Definition:**
Let $\mathbf{t}:=(t_1,\dots,t_n)$ be a vector of thresholds with $0<t_1<\dots<t_n<t_{n+1}:=+\infty$ and let $\boldsymbol\alpha:=(\alpha_1,\dots,\alpha_n)$ be a vector of Pareto alphas with $\alpha_i\ge 0$ and $\alpha_n>0$. The *piecewise Pareto* distribution} $\text{PPareto}(\mathbf{t},\boldsymbol\alpha)$ is defined by the distribution function $$
F_{\mathbf{t},\boldsymbol\alpha}(x):=\begin{cases}
0 & \text{ for $x<t_1$} \\
\displaystyle 1-\left(\frac{t_{k}}{x}\right)^{\alpha_k}\prod_{i=1}^{k-1}\left(\frac{t_i}{t_{i+1}}\right)^{\alpha_i} & \text{ for $x\in [t_k,t_{k+1}).$}
\end{cases}
$$

The family of piecewise Pareto distributions is very flexible:

**Proposition:** The set of Piecewise Pareto distributions is dense in the space of all positive-valued distributions (with respect to the Lévy metric).

This means that we can approximate any positive valued distribution as good as we want with piecewise Pareto. A very good approximation typically comes at the cost of many Pareto pieces. Piecewise Pareto  is often a good alternative to a discrete distribution, since it is much better to handle!

The Pareto package also provides functions for the picewise Pareto distribution. For instance:

*Layer mean:*
```{r}
t <- c(1000, 2000, 3000, 4000)
alpha <- c(2, 1, 3, 20)
PiecewisePareto_Layer_Mean(4000, 1000, t, alpha)
```

*Layer variance:*
```{r}
PiecewisePareto_Layer_Var(4000, 1000, t, alpha)
```

*Simulation:*
```{r}
rPiecewisePareto(20, t, alpha)
```

*CDF:*
```{r}
x <- c(1:10) * 1000
PiecewisePareto_CDF(x, t, alpha)
plot(PiecewisePareto_CDF(1:5000, t, alpha), xlab = "x", ylab = "CDF(x)")
```

*PDF:*
```{r}
PiecewisePareto_PDF(x, t, alpha)
plot(PiecewisePareto_PDF(1:5000, t, alpha), xlab = "x", ylab = "PDF(x)")
```

## PML Curves



## Matching a tower of layer losses


## References

Fackler, M. (2013) Reinventing Pareto: Fits for both small and large losses. ASTIN Colloquium Den Haag

Johnson, N.L., and Kotz, S. (1970) Continuous Univariate Distributions-I. Houghton Mifflin Co

Philbrick, S.W. (1985) A Practical Guide to the Single Parameter Pareto Distribution. PCAS LXXII: 44--84

Riegel, U. (2018) Matching tower information with piecewise Pareto. European Actuarial Journal 8(2): 437--460

Schmutz, M., and Doerr, R.R. (1998) Das Pareto-Modell in der Sach-Rueckversicherung. Formeln und Anwendungen. Swiss Re Publications, Zuerich

